---
title: "Lessons for Exploring Data: An application with LGU plant development data"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, warning = F, message = F}
library(learnr)
library(tidyverse)
knitr::opts_chunk$set(echo = FALSE,
                      warning = F, message = F)
lgu <- readRDS("data/lgu.rds")
rodents <- read_csv("data/portal_data_joined.csv")
```


## Reading in data

### Reading in an RDS

Write the R code required to read in our lgu data frame. For this exercise, pretend as if you are in your home directory and this data is stored in a folder called "data", located in your home directory. Call this data frame `lgu`.

```{r readrds, exercise=TRUE}

```

```{r readrds-hint}
lgu <- readRDS("data/lgu.rds")
```


### Reading in a CSV

Write the R code required to read in the rodents CSV. For this exercise, pretend as if you are in your home directory and this data is stored in a folder called "data", located in your home directory. Call this data frame `rodents``.

```{r readcsv, exercise=TRUE}

```

```{r readcsv-hint}
rodents <- read_csv("data/portal_data_joined.csv")
```


## Exploring data

### Data summaries 

Take a look at the LGU data. What kinds of data are in there? Use a number of summary functions (e.g. `summary()`, `str()`, `colnames()`) to explore. Annotate in your code with what you see. 

```{r explore, exercise=TRUE}







```

```{r explore-hint}
# You can experiment with a number of functions, such as the following:
## look at the column names
colnames(lgu)
## look at the structure with 
str(lgu)
## look at classes of individual columns
class(lgu$effective_date) # weird this isn't a number, right?
class(lgu$local_sales_number)
class(lgu$crop_name_common)
```

### Identifying NAs

Where are there NAs in the lgu data? Use the `is.na()` function in different columns to explore this

```{r explorena, exercise=TRUE}







```

```{r explorena-hint}
colnames(lgu)
table(is.na(lgu$ornamental))
table(is.na(lgu$crop_name_common))
table(is.na(lgu$name))
table(is.na(lgu$local_sales_number))
```

## Manipulating data

### Filtering data  
Let's try to summarize where our data is from. What states did we collect? Use the table() and/or unique() function to take a look at all of our states. Then let's subset our states to those for which we have company data. How does this change our states?

```{r states, exercise=TRUE}




```

```{r states-hint}
table(lgu$uni_states)

lgu_co <- lgu %>% 
  filter(!is.na(name))

table(lgu_co$uni_states)
```

### Selecting data  