---
title: Intro to Rmd
subtitle: And the YAML
author: Liza Wood
output: 
  html_document:
    toc: true
---


# Header 1

## How markdown works

**Bold font**

Rmarkdown docoument

-   Bulletpoints
    -   sub bullet point

[embedding links](www.google.com)

## How R works

```{r}
2+2 
```

# The YAML ^^  

YAMLs are the "set-up" for your Rmd file, indicated by three dashes to begin and end, and it can contain a series of information bits about the markdown document in between. Most important for us is the output, but this is a place to get more advanced (table of contents, LaTeX specification, ...).

What goes in a YAML also depends on the kind of output you have: PDF/LaTex, website, etc. See here:
https://cran.r-project.org/web/packages/ymlthis/vignettes/yaml-fieldguide.html

If you want to output as a pdf, you will need to set up LaTeX. (see more: https://bookdown.org/yihui/rmarkdown-cookbook/install-latex.html)


## YAML in R chunk

* echo = T: will include the code chunk
* include = F: will include the code chunk output or the code chunk
* results = c("asis"): asis is for LaTeX output
* fig.width = 7: default, but can adjust
* fig.height = 7: default, but can adjust
* warnings = F: prevent warnings from appearing

```{r, echo = F}
2+2
```

Use a setup chunk to set your default arguments inside your chunks

```{r setup, include = F}

# set default chunk options
knitr::opts_chunk$set(echo = T, warning = F, message = F, results = F)
# set default knit options
path_to_box <- '~/Box/lgu/'
knitr::opts_knit$set(root.dir = path_to_box)
```


```{r, results = T}
awards <- read.csv('data_clean/awards_selected_states.csv')
head(awards)
library(DT)
datatable(head(awards))
# OR -- these are the same, depending on whether you want to load in the library all together or just get one function
DT::datatable(head(awards))
```


```{r, message = T, warning = T}
library(sna)
library(igraph)
?degree
sna::degree
select
library(tidyverse)
```

