---
title: "Plant variety licensing descriptions"
author: "Liza Wood"
date: "2022-09-26"
output: html_document
---
```{r, warning= F, message=F}
library(tidyverse)
df <- readRDS("~/Box/lgu/data_clean/df_final.RDS")
pvp <- read.csv("~/Box/lgu/data_clean/pvpo_lgus.csv")
pto <- read.csv("~/Box/lgu/data_clean/uspto_lgus.csv")
#ip <- join pvp and pto, filter to 1980-2020
```

This page summarizes plant breeding activities at US Land Grant Universities (LGUs) and the subsequent trends in technology transfer between. Data for these summaries are combined across X sources:  
1. PVP database
1. USPTO database  
1. Licensing metadata from 25 LGUs, collected via FOIA requests

### How have LGUs been protecting their plant material?   

The PVP and USPTO databases were scraped and automated text matching was used to identify applications by Land Grant Universities. For PVPs, this meant looking for applications with Universities as Applicants, and for USPTO, this meant looking for LGUs as Assignees (and others?). Between 1980-2020 `r nrow(ip)` applications have been submitted by LGUs, X as PVP and X as USPTO. There have a steady increase in both IP types:

```{r}
# plot over time, with the two types of protections being charted
ggplot(ip) +
  geom_bar(aes(x = year), stat = "count")
```


### What plant material have LGUs been developing? 

For each of these protections, the protected crop was identified and categorized into one of 10 (?) groupings designated by the [FAO Agricultural Census](https://www.fao.org/3/a0135e/A0135E10.htm#app3), in line with Pardey et a. 20XX. The following crops have been   

```{r}
## A plot over time, colored by crop type
```


## What does licensing look like?  

Preliminary look at these data span the following features:  
* `r nrow(df)` licensing agreements across `r length(unique(df$uni_state))` Land Grant Universities between 2000-2020  
* Agreements are with `r length(unique(df$licensee[df$licensee_type == "company"]))` companies and `r length(unique(df$licensee[df$licensee_type == "other"]))` other entities (organizations, universities, individuals, etc.)  
* `r length(unique(df$variety_name))` unique plant varieties and/or breeding lines are licensed  
* These plant varieties span `r length(unique(df$crop_cat))` different crop categories, including: `r unique(df$crop_cat)`

```{r, echo = F, eval = F}
unique(df$license_yr)
df$yr_bi <- ifelse(df$license_yr %in% 2001:2010, "2001-2010", "2011-2020")
table(df$agreement_type)
618/(618+709) # Exclusive
table(df$agreement_type, df$yr_bi)
279/(279+234) # exclusive 2001-2010
339/(339+475) # exclusive 2011-2020
table(df$spatial_match)
prop.table(table(df$spatial_match))
381/(381+1434) # outside
prop.table(table(df$spatial_match, df$yr_bi), margin = 2)
158/(158+451) # outside 2001-2010
223/(223+983) # outside 2011-2020


df %>% 
  group_by(yr_bi) %>% 
  count()

table(df$company_size)
prop.table(table(df$company_size[df$company_size != "Unknown" & df$company_size != "Non-company"]))
prop.table(table(df$company_size, df$spatial_match), margin = 2)

df %>% 
  select(licensee, company_size) %>% 
  unique() %>% 
  count(company_size) %>% 
  mutate(sum = sum(n)) %>% 
  mutate(percent = 100*(n/sum)) %>% 
  filter(!is.na(company_size)) %>% 
  ggplot(aes(x = company_size, y = n)) +
  geom_bar(stat = "identity")
```

## Summaries over time   

```{r, out.width = "80%", fig.align= "center"}
df %>% 
  group_by(license_yr) %>% 
  count() %>% 
  ggplot(aes(x = license_yr, y = n,)) +
  geom_bar(stat = "identity") +
  stat_identity(geom = "text", colour = "white", size = 2,
                aes(label = n),
                position=position_stack(vjust=0.5)) +
  labs(x = "Year", y = "Count", title = "Licenses over time") +
  theme_linedraw() +
  theme(plot.title = element_text(hjust = 0.5, size = 10))
```

```{r, out.width = "80%", fig.align= "center"}
yr_count <- df %>% 
  group_by(license_yr) %>% 
  count() %>% 
  rename(yr_n = n)
colors9 <- RColorBrewer::brewer.pal(name = "Spectral", n = 9)

crop_plot <- df %>% 
  group_by(license_yr, crop_cat) %>% 
  count() %>% 
  left_join(yr_count) %>% 
  mutate(prop = round(100*n/sum(yr_n), 0)) %>% 
  ungroup() %>% 
  #mutate(prop = case_when(
  #  prop == 29 & license_yr == 2009 ~ 28,
  #  prop == 39 & license_yr == 2014 ~ 38,
  #  prop == 55 & license_yr == 2015 ~ 56,
  #  T ~ prop)) %>% 
  ggplot(aes(x = license_yr, y = prop, fill = crop_cat)) +
  geom_bar(stat = "identity") +
  #stat_identity(geom = "text", colour = "white", size = 3.5,
  #           aes(label = prop),
  #           position=position_stack(vjust=0.5)) +
  labs(x = "Year", y = "Percent", fill = "Crop",
       title = "Percent of licenses by crop category over time") +
  scale_fill_brewer(palette = "Spectral")+
  theme_linedraw() +
  theme(plot.title = element_text(hjust = 0.1, size = 10)) ; crop_plot
```

```{r, out.width = "80%", fig.align= "center"}
yr_count2 <- df %>% 
  filter(!(company_size %in% c("Unknown", "Non-company"))) %>% 
  filter(!is.na(company_size)) %>% 
  group_by(license_yr) %>% 
  count() %>% 
  rename(yr_n = n)

company_plot <- df %>% 
  filter(!(company_size %in% c("Unknown", "Non-company"))) %>% 
  filter(!is.na(company_size)) %>% 
  group_by(license_yr, company_size) %>% 
  count() %>% 
  left_join(yr_count2) %>% 
  mutate(prop = round(100*n/sum(yr_n), 0)) %>% 
  ungroup() %>% 
  #mutate(prop = case_when(
  #  prop == 48 & license_yr %in% c(2000, 2004) ~ 49,
  #  prop == 37 & license_yr == 2005 ~ 36,
  #  prop == 33 & license_yr == 2010 ~ 32,
  #  prop == 35 & license_yr == 2012 ~ 34,
  #  prop == 12 & license_yr == 2013 ~ 13,
  #  prop == 18 & license_yr == 2015 ~ 17,
  #  prop == 21 & license_yr == 2018 ~ 22,
  #  T ~ prop)) %>% 
  ggplot(aes(x = license_yr, y = prop, fill = company_size)) +
  geom_bar(stat = "identity") +
  #stat_identity(geom = "text", colour = "white", size = 3.5,
  #              aes(label = prop),
  #              position=position_stack(vjust=0.5)) +
  labs(x = "Year", y = "Percent", fill = "Company size",
       title = "Percent of licenses by company size over time") +
  scale_fill_manual(values = colors9[c(1,3,7,9)])+
  theme_linedraw() +
  theme(plot.title = element_text(hjust = 0.1, size = 10)) ; company_plot
```

```{r, out.width = "80%", fig.align= "center"}
licensee_plot <- df %>% 
  mutate(spatial_match = case_when(
    spatial_match == "Outside region" ~ "Extra-regional",
    spatial_match == "Within region" ~ "Regional",
    T ~ spatial_match
  )) %>% 
  group_by(license_yr, spatial_match) %>% 
  count() %>% 
  left_join(yr_count) %>% 
  mutate(prop = round(100*n/sum(yr_n), 0)) %>% 
  ggplot(aes(x = license_yr, y = prop, fill = spatial_match)) +
  geom_bar(stat = "identity") +
  stat_identity(geom = "text", colour = "white", size = 3.5,
                aes(label = prop),
                position=position_stack(vjust=0.5)) +
  labs(x = "Year", y = "Percent", fill = "Licensee location",
       title = "Percent of licenses by licensee over time") +
  scale_fill_manual(values = colors9[c(2,3,8)])+
  theme_linedraw() +
  theme(plot.title = element_text(hjust = 0.5, size = 10)) ; licensee_plot

```

```{r, out.width = "80%", fig.align= "center"}
yr_count3 <- df %>% 
  filter(!is.na(agreement_type)) %>% 
  group_by(license_yr) %>% 
  count() %>% 
  rename(yr_n = n)
agreement_plot <- df %>% 
  filter(!is.na(agreement_type)) %>% 
  group_by(license_yr, agreement_type) %>% 
  count() %>% 
  left_join(yr_count3) %>% 
  mutate(prop = round(100*n/sum(yr_n), 0)) %>% 
  ggplot(aes(x = license_yr, y = prop, fill = agreement_type)) +
  geom_bar(stat = "identity") +
  #stat_identity(geom = "text", colour = "white", size = 3.5,
  #              aes(label = prop),
  #              position=position_stack(vjust=0.5)) +
  labs(x = "Year", y = "Percent", fill = "Agreement restriction",
       title = "Percent of license by agreement type over time") +
  scale_fill_manual(values = colors9[c(1,3,8)])+
  theme_linedraw() +
  theme(plot.title = element_text(hjust = 0.1, size = 10)) ; agreement_plot
```

```{r, out.width = "80%", fig.align= "center"}
ggplot(df, aes(x = license_yr)) +
  geom_bar() +
  #geom_text(tab, aes(x = x, label = n)) + 
  labs(x = "Year", y = "Count", title = "Licenses over time by state") +
  theme_linedraw() +
  facet_wrap(~uni_state, nrow = 5) +
  coord_cartesian(ylim = c(0,150))
```

```{r, echo = F, eval = F}
# By crop type?
table(df$crop, df$license_yr)

df %>% 
  count(license_yr, crop_fewer) %>% 
  group_by(license_yr) %>% 
  mutate(sum = sum(n)) %>% 
  mutate(percent = 100*(n/sum)) %>% 
  ggplot(aes(x = license_yr, y = percent, fill = crop_fewer)) +
  geom_bar(stat = "identity") +
  labs(x = "Year", y = "Percent", 
       fill = "Crop type",
       title = "Licenses over time by crop") +
  #scale_fill_viridis_d() +
  scale_fill_brewer(palette = "Spectral") +
  theme_linedraw()

df %>% 
  count(license_yr, crop_intermed) %>% 
  group_by(license_yr) %>% 
  mutate(sum = sum(n)) %>% 
  mutate(percent = 100*(n/sum)) %>% 
  ggplot(aes(x = license_yr, y = percent)) +
  geom_bar(stat = "identity") +
  geom_smooth() +
  labs(x = "Year", y = "Percent", 
       fill = "Crop type",
       title = "Licenses over time by crop") +
  #scale_fill_viridis_d() +
  theme_linedraw() +
  facet_wrap(~crop_intermed, scales = "free_y")

df %>% 
  count(state, license_yr, crop_fewer) %>% 
  group_by(state, license_yr) %>% 
  mutate(sum = sum(n)) %>% 
  mutate(percent = 100*(n/sum)) %>% 
  ggplot(aes(x = license_yr, y = percent, fill = crop_fewer)) +
  geom_bar(stat = "identity") +
  labs(x = "Year", y = "Percent", 
       fill = "Crop type",
       title = "Licenses over time by crop and state") +
  #scale_fill_viridis_d() +
  scale_fill_brewer(palette = "Spectral") +
  theme_linedraw() +
  facet_wrap(~state, nrow = 3)

# By exclusivity type?
table(df$agreement_type, df$license_yr) 

df %>% 
  count(license_yr, agreement_type) %>% 
  group_by(license_yr) %>% 
  mutate(sum = sum(n)) %>% 
  mutate(percent = 100*(n/sum)) %>% 
  ggplot(aes(x = license_yr, y = percent, fill = agreement_type)) + 
  geom_bar(stat = "identity") +
  labs(x = "Year", y = "Percent", 
       fill = "Agreement type",
       title = "Licenses over time by license agreement") +
  scale_fill_brewer(palette = "Spectral") +
  theme_linedraw()

df %>% 
  count(state, license_yr, agreement_type) %>% 
  group_by(state, license_yr) %>% 
  mutate(sum = sum(n)) %>% 
  mutate(percent = 100*(n/sum)) %>% 
  ggplot(aes(x = license_yr, y = percent, fill = agreement_type)) + 
  geom_bar(stat = "identity") +
  facet_wrap(~state, nrow = 3) +
  labs(x = "Year", y = "Percent", 
       fill = "Agreement type",
       title = "Licenses over time by license agreement and state") +
  scale_fill_brewer(palette = "Spectral") +
  theme_linedraw()

# By companies versus non companies?
df %>% 
  filter(!is.na(licensee_type)) %>% 
  count(license_yr, licensee_type) %>% 
  group_by(license_yr) %>% 
  mutate(sum = sum(n)) %>% 
  mutate(percent = 100*(n/sum)) %>% 
  ggplot(aes(x = license_yr, y = percent, fill = licensee_type)) + 
  geom_bar(stat = "identity") +
  labs(x = "Year", y = "Percent", 
       fill = "Licensee",
       title = "Licenses over time by licensee") +
  scale_fill_viridis_d() +
  theme_linedraw()

df %>% 
  filter(!is.na(licensee_type)) %>% 
  count(state, license_yr, licensee_type) %>% 
  group_by(state, license_yr) %>% 
  mutate(sum = sum(n)) %>% 
  mutate(percent = 100*(n/sum)) %>% 
  ggplot(aes(x = license_yr, y = percent, fill = licensee_type)) + 
  geom_bar(stat = "identity") +
  facet_wrap(~state, nrow = 3) +
  labs(x = "Year", y = "Percent", 
       fill = "Licensee",
       title = "Licenses over time by licensee and state") +
  scale_fill_viridis_d() +
  theme_linedraw()

# By company size?
df %>% 
  filter(company_size != "Unknown") %>% 
  count(license_yr, company_size) %>% 
  group_by(license_yr) %>% 
  mutate(sum = sum(n)) %>% 
  mutate(percent = 100*(n/sum)) %>% 
  ggplot(aes(x = license_yr, y = percent, fill = company_size)) + 
  geom_bar(stat = "identity") +
  labs(x = "Year", y = "Percent", 
       fill = "Company size",
       title = "Licenses over time by company size") +
  scale_fill_brewer(palette = "Spectral") +
  theme_linedraw()

df %>% 
  filter(company_size != "Unknown") %>% 
  count(state, license_yr, company_size) %>% 
  group_by(state, license_yr) %>% 
  mutate(sum = sum(n)) %>% 
  mutate(percent = 100*(n/sum)) %>% 
  ggplot(aes(x = license_yr, y = percent, fill = company_size)) + 
  geom_bar(stat = "identity") +
  facet_wrap(~state, nrow = 3) +
  labs(x = "Year", y = "Percent", 
       fill = "Company size",
       title = "Licenses over time by company size and state") +
  scale_fill_brewer(palette = "Spectral") +
  theme_linedraw()

df %>% 
  filter(company_size != "Unknown") %>% 
  ggplot(aes(x = factor(license_yr), y = rev_log)) + 
  geom_boxplot() 

# Within and outside of region
colors <- RColorBrewer::brewer.pal(name = "Spectral", n = 9)[c(3,8)]
df %>% 
  filter(spatial_match != "Unknown") %>% 
  mutate(spatial_match = case_when(
    domestic == F ~ "International",
    T ~ spatial_match)) %>% 
  count(license_yr, spatial_match) %>% 
  group_by(license_yr) %>% 
  mutate(sum = sum(n)) %>% 
  mutate(percent = 100*(n/sum)) %>% 
  ggplot(aes(x = license_yr, y = percent, fill = spatial_match)) + 
  geom_bar(stat = "identity") +
  labs(x = "Year", y = "Percent", 
       fill = "Licensee location",
       title = "Licenses over time by licensee geographic scope") +
  scale_fill_manual(values = colors) +
  theme_linedraw()


df %>% 
  filter(spatial_match != "Unknown") %>% 
  mutate(spatial_match = case_when(
    domestic == F ~ "International",
    T ~ spatial_match)) %>% 
  count(state, license_yr, spatial_match) %>% 
  group_by(state, license_yr) %>% 
  mutate(sum = sum(n)) %>% 
  mutate(percent = 100*(n/sum)) %>% 
  ggplot(aes(x = license_yr, y = percent, fill = spatial_match)) + 
  geom_bar(stat = "identity") + 
  facet_wrap(~ state) +
  labs(x = "Year", y = "Percent", 
       fill = "Licensee location",
       title = "Licenses over time by licensee geographic scope and state") +
  scale_fill_manual(values = colors) +
  theme_linedraw()
```

## Summaries by crop   

```{r, out.width = "80%", fig.align= "center"}
crop_count <- df %>% 
  group_by(crop_cat) %>% 
  count() %>% 
  arrange(desc(n))
df$crop_cat <- factor(df$crop_cat, levels = crop_count$crop_cat)

ggplot(df, aes(x = crop_cat)) +
  geom_bar() +
  stat_count(geom = "text", colour = "white", size = 3.5,
             aes(label = ..count..),
             position=position_stack(vjust=0.5)) +
  labs(x = "Crop", y = "Count", 
       title = "Licenses by crop") +
  theme_linedraw() +
  theme(axis.text.x= element_text(angle = 45, hjust = 1)) 
```

```{r, eval = F, echo = F}
crop_by_state <- data.frame(table(df$crop_cat, df$state))

ggplot(df, aes(x = crop_cat)) +
  geom_bar() +
  labs(x = "Crop", y = "Count", title = "License totals by crop and state") +
  theme_linedraw() +
  facet_wrap(~state, nrow = 3) +
  theme(axis.text.x= element_text(angle = 45, hjust = 1))
```

## What are the innovative qualities for food crops?  


```{r, eval = F, echo = F}
# By exclusivity type?
#crop_by_agreement <- data.frame(table(df$crop_cat, df$agreement_type))

df %>% 
  count(crop_cat, agreement_type) %>% 
  group_by(crop_cat) %>% 
  mutate(sum = sum(n)) %>% 
  mutate(percent = 100*(n/sum)) %>% 
  ggplot(aes(x = crop_cat, y = percent, fill = agreement_type)) + 
  geom_bar(stat = "identity") +
  labs(x = "Crop", y = "Percent", 
       fill = "Agreement type",
       title = "Crop licenses by agreement type") +
  scale_fill_brewer(palette = "Spectral") +
  theme_linedraw() +
  theme(axis.text.x= element_text(angle = 45, hjust = 1))

df %>% 
  count(state, crop, agreement_type) %>% 
  group_by(state, crop) %>% 
  mutate(sum = sum(n)) %>% 
  mutate(percent = 100*(n/sum)) %>% 
  ggplot(aes(x = crop, y = percent, fill = agreement_type)) + 
  geom_bar(stat = "identity") +
  labs(x = "Crop", y = "Percent", 
       fill = "Agreement type",
       title = "Crop licenses by agreement type and state") +
  scale_fill_viridis_d() +
  theme_linedraw() +
  theme(axis.text.x= element_text(angle = 45, hjust = 1)) +
  facet_wrap(~state, nrow = 3)

# By companies versus non companies?

crop_by_licenseetype <- data.frame(table(df$crop_cat, df$licensee_type))

df %>% 
  filter(!is.na(licensee_type)) %>% 
  count(crop_cat, licensee_type) %>% 
  group_by(crop_cat) %>% 
  mutate(sum = sum(n)) %>% 
  mutate(percent = 100*(n/sum)) %>% 
  ggplot(aes(x = crop_cat, y = percent, fill = licensee_type)) + 
  geom_bar(stat = "identity") +
  labs(x = "Crop", y = "Percent", 
       fill = "Licensee",
       title = "Crop licenses by licensee") +
  scale_fill_manual(values = colors) +
  theme_linedraw() +
  theme(axis.text.x= element_text(angle = 45, hjust = 1))

df %>% 
  filter(!is.na(licensee_type)) %>% 
  count(state, crop_cat, licensee_type) %>% 
  group_by(state, crop_cat) %>% 
  mutate(sum = sum(n)) %>% 
  mutate(percent = 100*(n/sum)) %>% 
  ggplot(aes(x = crop_cat, y = percent, fill = licensee_type)) + 
  geom_bar(stat = "identity") +
  labs(x = "Crop", y = "Percent", 
       fill = "Licensee",
       title = "Crop licenses by licensee and state") +
  scale_fill_manual(values = colors) +
  theme_linedraw() +
  theme(axis.text.x= element_text(angle = 45, hjust = 1)) +
  facet_wrap(~state, nrow = 3)

# By company size?

crop_by_companysize <- data.frame(table(df$crop, df$company_size))

df %>% 
  filter(company_size != "Unknown") %>% 
  count(crop_cat, company_size) %>% 
  group_by(crop_cat) %>% 
  mutate(sum = sum(n)) %>% 
  mutate(percent = 100*(n/sum)) %>% 
  ggplot(aes(x = crop_cat, y = percent, fill = company_size)) + 
  geom_bar(stat = "identity") +
  labs(x = "Year", y = "Percent", 
       fill = "Company size",
       title = "Crop licenses by company size") +
  scale_fill_brewer(palette =  "Spectral") +
  theme_linedraw() +
  theme(axis.text.x= element_text(angle = 45, hjust = 1))

df %>% 
  filter(company_size != "Unknown") %>% 
  count(state, crop_cat, company_size) %>% 
  group_by(state, crop_cat) %>% 
  mutate(sum = sum(n)) %>% 
  mutate(percent = 100*(n/sum)) %>% 
  ggplot(aes(x = crop_cat, y = percent, fill = company_size)) + 
  geom_bar(stat = "identity") +
  labs(x = "Year", y = "Percent", 
       fill = "Company size",
       title = "Crop licenses by company size and state") +
  scale_fill_brewer(palette =  "Spectral") +
  theme_linedraw() +
  theme(axis.text.x= element_text(angle = 45, hjust = 1)) +
  facet_wrap(~state, nrow = 3)

# Within and outside of region
crop_by_companylocation <- data.frame(table(df$crop_cat, df$inregion))

df %>% 
  filter(spatial_match != "Unknown") %>% 
  mutate(spatial_match = case_when(
    domestic == F ~ "International",
    T ~ spatial_match)) %>% 
  count(crop_cat, spatial_match) %>% 
  group_by(crop_cat) %>% 
  mutate(sum = sum(n)) %>% 
  mutate(percent = 100*(n/sum)) %>% 
  ggplot(aes(x = crop_cat, y = percent, fill = spatial_match)) + 
  geom_bar(stat = "identity") +
  labs(x = "Year", y = "Percent", 
       fill = "Licensee location",
       title = "Crop licenses by licensee geographic scope") +
  scale_fill_manual(values = colors) +
  theme_linedraw() +
  theme(axis.text.x= element_text(angle = 45, hjust = 1))

df %>% 
  filter(spatial_match != "Unknown") %>% 
  mutate(spatial_match = case_when(
    domestic == F ~ "International",
    T ~ spatial_match)) %>% 
  count(state, crop, spatial_match) %>% 
  group_by(state, crop) %>% 
  mutate(sum = sum(n)) %>% 
  mutate(percent = 100*(n/sum)) %>% 
  ggplot(aes(x = crop, y = percent, fill = spatial_match)) + 
  geom_bar(stat = "identity") +
  labs(x = "Year", y = "Percent", 
       fill = "Licensee location",
       title = "Crop licenses by licensee geographic scope and state") +
  scale_fill_viridis_d() +
  theme_linedraw() +
  theme(axis.text.x= element_text(angle = 45, hjust = 1)) +
  facet_wrap(~state, nrow = 3)


# ---- Combining with USDA funding data ----

# Trends in federal money contribution by state and licensing?

funding %>% 
  ggplot(aes(x = year, y = Award.Dollars/1000000)) +
  geom_bar(stat = "identity") +
  labs(x = "Year", y = "Public award dollars (1M)", 
       title = "USDA plant breeding funding (All 38 funded states)") +
  theme_linedraw() 

fund_df %>% 
  ggplot(aes(x = funding_yr_grpd, y = funding_amt_grpd/1000000)) +
  geom_bar(stat = "identity") +
  labs(x = "Year", y = "Public award dollars (1M)", 
       title = "USDA plant breeding funding by state") +
  theme_linedraw() +
  facet_wrap(~state, nrow = 5) +
  theme(axis.text.x= element_text(angle = 45, hjust = 1))

# This is different because this is only for the schools we have
funding %>% 
  filter(state %in% df$state) %>% 
  group_by(state) %>% 
  summarize(sum(Award.Dollars))

funding %>% 
  filter(state %in% df$state) %>% 
  ggplot(aes(x = year, y = Award.Dollars/1000000)) +
  geom_bar(stat = "identity") +
  labs(x = "Year", y = "Public award dollars (1M)", 
       title = "USDA plant breeding funding (14 states)") +
  theme_linedraw() 

df %>% 
  filter(!is.na(funding_yr_grpd)) %>% 
  select(state, funding_yr_grpd, funding_amt_grpd) %>% 
  unique() %>% 
  ggplot(aes(x = funding_yr_grpd, y = funding_amt_grpd/1000000)) +
  geom_bar(stat = "identity") +
  labs(x = "Year", y = "Public award dollars (1M)", 
       title = "USDA Plant Breeding Funding for states in our sample") +
  theme_linedraw() +
  facet_wrap(~state, nrow = 3) +
  theme(axis.text.x= element_text(angle = 45, hjust = 1))


# Exclusivity ~ crop_cat + company_revenue + (1 | grp_yr_license) + grp_amt
library(lme4)
factor(df$crop, levels = c("Field crops", ))
lm(rev_log ~ crop, df)

ex <- df %>% select(agreement_bi, funding_amt_grpd, rev) %>% 
  filter(agreement_bi == 1)
nex <- df %>% select(agreement_bi, funding_amt_grpd, rev) %>% 
  filter(agreement_bi == 0)

t.test(ex$funding_amt_grpd, nex$funding_amt_grpd)
# the mean of public funding is significantly higher for non-exclusive licenses than for exclusive licenses
t.test(ex$rev, nex$rev)
# the mean of licensee revenue is not significantly higher, though higher, for non-exclusive licenses than for exclusive licenses
df$agreement_bi_f <- factor(df$agreement_bi)
df$agreement_bi_f <- recode_factor(df$agreement_bi_f, '0' = "Non-exclusive", '1' = "Exclusive")

df %>% 
  filter(!is.na(agreement_bi_f)) %>% 
  ggplot(aes(x = agreement_bi_f, 
             y = funding_amt_grpd/1000000, 
             color = factor(agreement_bi_f))) +
  geom_boxplot() +
  scale_color_viridis_d() +
  theme_linedraw() +
  labs(x = "", y = "Public award dollars (1M)", color = "Agreement type")

df %>% 
  filter(!is.na(agreement_bi_f)) %>% 
  ggplot(aes(x = agreement_bi_f, 
             y = rev/1000000, 
             color = factor(agreement_bi_f))) +
  geom_boxplot() +
  scale_color_viridis_d() +
  theme_linedraw() +
  labs(x = "", y = "Licensee annual revenue (1M)", color = "Agreement type") +
  ylim(0,200)

inr <- df %>% select(inregion, funding_amt_grpd, rev) %>% 
  filter(inregion == 1)
outr <- df %>% select(inregion, funding_amt_grpd, rev) %>% 
  filter(inregion == 0)

t.test(inr$funding_amt_grpd, outr$funding_amt_grpd)
# the mean of public funding is significantly higher for out of region licenses than for in region licenses
t.test(inr$rev, outr$rev)
# the mean of licensee revenue is not significantly higher, though higher, for non-exclusive licenses than for exclusive licenses
df$inregion_f <- factor(df$inregion)
df$inregion_f <- recode_factor(df$inregion_f, '0' = "Outside region", '1' = "Within region")

df %>% 
  filter(!is.na(inregion_f)) %>% 
  ggplot(aes(x = inregion_f, 
             y = funding_amt_grpd/1000000, 
             color = factor(inregion_f))) +
  geom_boxplot() +
  scale_color_viridis_d() +
  theme_linedraw() +
  labs(x = "", y = "Licensee annual revenue (1M)", color = "Region of licensee")

df %>% 
  filter(!is.na(inregion_f)) %>% 
  ggplot(aes(x = inregion_f, 
             y = rev/1000000, 
             color = factor(inregion_f))) +
  geom_boxplot() +
  scale_color_viridis_d() +
  theme_linedraw() +
  labs(x = "", y = "Public award dollars (1M)", color = "Region of licensee")

inc <- df %>% select(domestic, funding_amt_grpd, rev) %>% 
  filter(domestic == T)
outc <- df %>% select(domestic, funding_amt_grpd, rev) %>% 
  filter(domestic == F)

t.test(inc$funding_amt_grpd, outc$funding_amt_grpd)
# the mean of public funding is significantly higher for out of country licenses than for out of country licenses
t.test(inc$rev, outc$rev)
# the mean of licensee revenue is not significantly higher, though higher, for non-exclusive licenses than for exclusive licenses


df %>% 
  filter(!is.na(inregion_f)) %>% 
  ggplot(aes(x = factor(license_yr), 
             y = rev/1000000)) +
  geom_boxplot() +
  scale_color_viridis_d() +
  theme_linedraw() +
  labs(x = "", y = "Licensee annual revenue (1M)") +
  ylim(0,2000)

summary(lm(license_yr ~ rev + funding_amt_grpd, data = df))

nlicenses <- df %>% 
  group_by(funding_yr_grpd, state) %>% 
  count()

bystate <- df %>% 
  group_by(state, funding_yr_grpd, funding_amt_grpd_log) %>% 
  summarize(avg_rev = mean(rev, na.rm = T)) %>% 
  left_join(nlicenses)

summary(lm(n ~ funding_amt_grpd_log + avg_rev, data = bystate))

ggplot(bystate, aes(x = funding_amt_grpd_log, y = n)) +
  geom_point() +
  geom_smooth(method = 'lm')

df.full <- df %>% 
  select(crop, funding_amt_grpd_log, license_yr_grpd, state) %>% 
  filter(complete.cases(.) == T)
multi <- nnet::multinom(factor(crop) ~ rev_log + funding_amt_grpd_log, data = df, Hess = T)
summary(multi)


df.full <- df %>% 
  select(agreement_bi, crop, rev_log, funding_amt_grpd_log, license_yr_grpd, state) %>% 
  filter(complete.cases(.) == T)

glm_agree <- glm(agreement_bi ~ crop + rev_log + funding_amt_grpd_log, 
                 family = binomial,  data = df.full)
summary(glm_agree)

glmer_agree <- glmer(agreement_bi ~ crop + rev_log + funding_amt_grpd_log + 
              (1 | state), #  (1|license_yr_grpd) + 
             family = binomial,  data = df.full)
summary(glmer_agree)

df.full <- df %>% 
  select(inregion, crop, rev_log, funding_amt_grpd_log, license_yr_grpd, state) %>% 
  filter(complete.cases(.) == T)

glm_inregion <- glm(inregion ~ crop + rev_log + funding_amt_grpd_log, 
                 family = binomial,  data = df.full)
summary(glm_inregion)

glmer_inregion <- glmer(inregion ~ crop + rev_log + funding_amt_grpd_log + 
                (1|state), # (1|license_yr_grpd)
             family = binomial,  data = df.full)
summary(glmer_inregion)

```