```{r}
DT::datatable()
```


```{r}
library(tidyverse)
df3 <- read.csv("~/Box/lgu/data_clean/joined_inventions_licenses_awardstotal.csv")
```

* Note: Is the sum the best thing to do in the join to account for time? (DO THIS)
* Is there a relationship between company size and 'awards money in' to different varieties?  
* Is there a relationship between license type (Exclusive vs not) and 'awards money in' to different varieties?  
  * If the crop is sold non-exclusively, how many licensees are there? Can we get a sense of how much money in is then 'redistributed' (broad sense) back into licensees? 


Reverse engineer: get just each unique variety and their funding amount (funds associated with breeder) -- need to remove the company info, since that is the dupliate

```{r}
df2 <- df3 %>% 
  select(-c(licensee:size_range_sales)) %>% 
  unique()


```

What are the total funds per each crop category?

```{r}
df2 %>% 
  group_by(crop_cat) %>% 
  summarize(funds_per_crop = sum(total_avg_funds, na.rm = T)) %>% 
  arrange(-funds_per_crop)
```

What about average?

```{r}
df2 %>% 
  group_by(crop_cat) %>% 
  summarize(avg_funds_per_crop = mean(total_funds, na.rm = T)) %>% 
  arrange(-avg_funds_per_crop)
```


```{r}
dft <- read.csv("~/Box/lgu/data_clean/joined_inventions_licenses_awards_timeperiod.csv")
```

