# Who buys what?  

------
A summary of plant material licensing for LGU Agricultural Colleges, 2000-2020  
Authors: Adam Vera, Liza Wood, Victoria Fletcher  
Date: `r Sys.Date()`
------

```{r setup, include = F}
library(tidyverse)
library(kableExtra)
path_to_box <- '~/Box/lgu/'
knitr::opts_chunk$set(echo = F, warning = F, message = F, results = T, fig.align = 'center')
knitr::opts_knit$set(root.dir = path_to_box)
```

```{r, echo = F}
setwd(path_to_box)
licenses <- read.csv("data_clean/license_clean.csv")
companies <- read.csv("data_clean/company_db_full_clean.csv")
other_licensees <- read.csv("data_indices/other_licensee.csv")
```


## Introduction  

US Land Grant Universities (LGUs) are an important contributor to public plant breeding. Universities develop different plant material and varieties and then (increasingly) _license_ the plant material they develop to third parties to commercialize and/or continue to innovation. This report summarizes plant variety licenses between LGUs and the companies that purchase these products. This analysis explores the role specific universities have in this market, the breakdown of crops licensed by university, and the financial profile of the companies that purchase the licenses from LGUs. 

We base this report on data collected via public records requests. These records include licensing agreement data (what material is being licensed and to whom) between 2000-2020 for 25 LGUs (`r paste(unique(licenses$uni_state), collapse = ", ")`. We requested data from all 50 states, but only received data from half due to public records requests denials or prohibitive costs. 

Key take-aways are as follows:  

* `r nrow(licenses)` licensing agreements across `r length(unique(licenses$uni_state))` Land Grant Universities between 2000-2020  
* Agreements are with `r length(unique(companies$licensee[companies$licensee_type == "company"]))` companies and `r length(unique(other_licensees$official_name))` other entities (organizations, universities, individuals, etc.)  
* `r length(unique(licenses$variety_name))` unique plant varieties and/or breeding lines are licensed

## What crops are being licensed? 

```{r, echo=FALSE, eval=FALSE}
colnames(licenses)
```

<!----
## License data frame column names and descriptions

  1) X: Row Number
  2) crop_name_common: common name of each crop
  3) crop_name_scientific: scientific name of each crop with genus and species
  4) variety_name: name of specific variety given by the inventor
  5) licensee: buyer of license
  6) agreement_type: type of agreement between licensee and uni_state; for example: exclusive, non-exclusive, etc.
  7) effective_date: date of purchase/license
  8) end_date: termination date of license agreement
  9) uni_state: state of university
  10) ip_name: intellectual property crop name
  11) fao_class1: general classification of crop given by Food and Agricultural Organization (FAO)
  12) fao_class2: more specific classification of crop given by FAO
  13) ornamental: marks whether the crop is an ornamental (TRUE) or not (FALSE)
  14) orn_use: type of ornamental for such crops that are TRUE
  15) crop_cat: same as fao_class1 with proper formatting
--->
<!--- 2. Next we want to make some summaries of the licensing data (again, these may be similar to what we practiced last term). Take a look at the columns we have and think about what you'd like to know about the data. For instance: 
  * How many different licenses do we have...
    * Per year? --->
    
We categorize the crops LGUs are developing into one of 10 groupings designated by the [FAO Agricultural Census](https://www.fao.org/3/a0135e/A0135E10.htm#app3), and we can summarize these crop types across different universities in our data. Note that we remove ornamental plants from this analysis to focus in only on food and cash crops.  

The crops most often licensed resemble the general trend of what we see in LGU plant variety development protection: cereals, fruits and nuts, and oilseeds. The crop licensed the most of all crop categories is cereals, making up 46% of licenses. The next most popular categories are fruits and nuts (17%) and oilseeds (16%). Cereals in our licensing data are comprised of primarily wheat (73%), with barley a far second (18%), the remaining nine crops in the cereal category leaving a long tail (e.g. sorghum, oat, corn, rice, millet, teff).

<!----NAs values for FAO class mostly resemble ornamental crops and have been removed. This report is more focused on food and cash crops. -- worth realizing that NAs in the fao_class1 variable become ornamental in the crop_cat variable if ornamental varible = T--->


```{r, eval = F}
licenses %>% 
  filter(!is.na(fao_class1)) %>%
  count()
  
licenses %>% 
  filter(!is.na(fao_class1)) %>%
  group_by(fao_class1) %>%
  count() %>%
  arrange(-n)
```

```{r}
years <-  as.Date(licenses$effective_date, '%Y-%m-%d')

effective_year <- as.numeric(format((as.Date(licenses$effective_date, '%Y-%m-%d')),'%Y'))

licenses <- licenses %>%
  mutate(effective_year = as.numeric(format((as.Date(effective_date, '%Y-%m-%d')),'%Y'))) %>% 
    filter(effective_year <= 2020 & effective_year >= 2000) 
```

```{r}
# Just removing this from the start since we say we are only thinking about ornamental. This may change things
licenses <- licenses %>% 
  filter(crop_cat != "Ornamental") 

# re-writing with crop_cat variable because the text is cleaner
crop_table <- licenses %>% 
    filter(!is.na(crop_cat)) %>%
  group_by(crop_cat) %>%
  count() %>%
  arrange(-n) %>% 
  ungroup() %>% 
  mutate(percent = paste0(round(100*n/sum(n)), "%"))

kable(crop_table, col.names = c("Crop category", "Count of licenses", "Percent of all licenses")) %>% 
  kable_styling()
```

```{r, eval = F}
# I decided not to add this
licenses_cereals <- licenses %>% 
  filter(!is.na(crop_cat)) %>%
  filter(crop_cat == "Cereals", !is.na(crop_cat))

cereal_table <- licenses_cereals %>%
  group_by(crop_name_common) %>%
  count() %>%
  arrange(-n) %>% 
  ungroup() %>% 
  mutate(percent = paste0(round(100*n/sum(n)), "%"))

kable(cereal_table, col.names = c("Cereal type", "Count of licenses", "Percent of all licenses")) %>% 
  kable_styling()
```

Washington dominates the cereal licensing at 2122 licenses, licensing 65.6% (2122/3234) of cereals. The next largest producer of cereals is Oregon, at 15% (485/3234). In the table below, we calculate the count and percentage of crops in all of the states in our sample. 


```{r}
statecrop_table <- licenses %>% 
  filter(!is.na(crop_cat)) %>%
  group_by(crop_cat, uni_state) %>%
  count() %>%
  arrange(-n) %>% 
  ungroup() %>% 
  mutate(percent = paste0(round(100*n/sum(n)), "%"))
# Do I want to add a percentage here?

DT::datatable(statecrop_table, rownames = F,
              colnames = c("Crop category", "LGU", "Count of licenses", "Percent of all licenses"))
```

```{r, echo=FALSE, eval=FALSE}
licenses %>%
  group_by(uni_state, effective_year, year_range) %>%
  count() %>%
  ungroup() %>%
  group_by(uni_state, year_range) %>%
  summarize(mean = mean(n, na.rm = T))
```

## How has licensing changed over time? 

We look at the date of the licensing agreements to understand trends in university licensing over time. Across the 25 universities, licensing has been steadily increasing, which aligns with the trends of plant variety innovation generally (Section 1).


```{r, fig.align= "center", echo = F, out.width = "100%"}
licenses %>% 
  group_by(effective_year) %>% 
  count() %>% 
  ggplot(aes(x = effective_year, y = n)) +
  geom_bar(stat = "identity", fill = "#472D7BFF") +
  stat_identity(geom = "text", colour = "white", size = 3,
                aes(label = n),
                position=position_stack(vjust=0.5)) +
  labs(x = "", y = "Count", title = "Licenses over time") +
  theme_minimal()
```

```{r, echo = F}
counts <- licenses %>% group_by(uni_state) %>% count()
yrly_counts <- licenses %>% group_by(uni_state, effective_year) %>% count()
```

The amount of licensing varies by state considerably. The number of licenses over the twenty year periods= ranges from only `r min(counts$n)` in `r counts$uni_state[counts$n == min(counts$n)]` to `r max(counts$n)` in `r counts$uni_state[counts$n == max(counts$n)]`, with an annual average of `r round(mean(yrly_counts$n))` licenses. 
    
<!--- To accurately represent this, manipulations of the data frame is required to determine the effective year of the license as well as group licenses into five year periods.--->
```{r}

licenses <- licenses %>%
  filter(effective_year < 2021) %>%
  mutate(year_range = case_when(effective_year > 2015 ~ "2016-2020", 
                                effective_year > 2010 ~ "2011-2015", 
                                effective_year > 2005 ~ "2006-2010", 
                                effective_year >= 2000 ~ "2000-2005", 
                                TRUE ~ "Older")) %>% 
  # Liza adding in a filter to remove older ones because I only requested data from 2000 and so the pre-2000 data can't really be compared across states
  filter(!(year_range %in% c("1996-2000", "Older")))
``` 

```{r, echo= FALSE, eval = FALSE}
licenses %>% 
  group_by(effective_year) %>%
  count() %>%
  arrange(-n)

licenses %>% 
  group_by(year_range) %>%
  count() %>% 
  arrange(-n)
```

```{r, results = F}
licenses %>%
  group_by(uni_state) %>%
  count() %>%
  arrange(-n)

top_uni_list <- c("Washington", "Oregon", "North Carolina", "Texas", "Iowa")
```

To evaluate this variation across states, we group licenses into five-year time periods to present changes in licensing trends across different universities. Given our limited ability to observe trends across all 25 universities at once, we also choose to identify and highlight only the top five universities by license count (`r top_uni_list`), while the remaining twenty universities are considered to be "Other". This is represented in a plot below per five year time period.

```{r}
licenses <- licenses %>%
  mutate(top_uni = case_when(uni_state %in% top_uni_list ~ uni_state, TRUE ~ "Other"))

licenses$top_uni <- factor(licenses$top_uni, levels = c("Other", "Iowa", "Texas", "North Carolina", "Oregon", "Washington"))

ggplot(licenses, aes(x = factor(year_range, 
                                level=c('2000-2005', '2006-2010', '2011-2015', '2016-2020')),
                     fill = top_uni)) + 
  geom_bar() +
  theme(axis.text.x = element_text(angle=70, vjust=0.5)) +
  scale_fill_viridis_d(option = "D") +
  labs(title = "License count by university",
       x = "Year Range", y = "License Count", fill = "University") +
  theme_minimal()
```

## Who is licensing LGU plant material?  

<!--- 3. What are the columns in our companies data? Write out some 'data documentation' by making a list of all the columns and writing our short descriptions of each of them. These will help you and your collaborators know what they're looking at when they first access the data. If there are any irrelevant columns, remove them with the select function. --->

```{r, results = F}
colnames(companies)[2] <- "dnb_name"
companies <- companies[-c(10)]
colnames(companies)
```

<!---
## Column data frame column names and descriptions

  1) X: row number
  2) dnb_name: company name (from dnb)
  3) company_type: type of company
  4) website: url to company website
  5) parent: parent company if part of a larger corporation
  6) corp_link: number of companies under parent?
  7) industry: type of business
  8) currency: currency used by business and of assets
  9) assets: wealth
  10) address: business address
  11) naics2017: North American Industry Classification System
  12) yr_founded: year founded
  13) incorp_location: Additional location
  14) local_employee_number: number of employees at business location on license
  15) total_employee_number: domestic employee number
  16) global_employee_number: international employee number
  17) local_sales_number: gross income of business at location on license
  18) global_sales_number: total gross income of company
  19) db_type: database name
  20) licensee: official name from master company list on licenses spreadsheet (used for joining)
  21) zipcode: zip code
  22) lat: latitude coordinate of company location
  23) long: longitude coordinate of company location
  24) country: country of company HQ location
  25) state: US state of HQ
  26) domestic: logical to determine if international or not; TRUE means US based
  27) licensee_type: company or other
  28) intermediate_employee_number: irrelevant column for calculations only
  29) highest_employee_number: largest available data points for company employee number
  30) highest_sales_number: largest available data points for company sales money
  31) size_range_sales: determination of company size based on sales number
  32) size_range_employee: determination of company size based on employee number
  
  Two data frames are joined for this report: Companies and Licenses. Licenses describe the agreement between universities and companies, while the Companies include information regarding the companies themselves. The unique number of licensees in the licenses data frame is larger than the number of companies in the companies data frame. This suggests that not all licensees are documented by the companies database used in this report. This is because some licensees are listed as individuals and various associations that are not considered to be companies.
--->

Our team has used the licensing records to build out a data base of companies and organizations who have licensed material from LGUs. We have relied on the D&B Hoovers data base for most of this company data, including information such as the location of company headquarters and their size, as measured by sales (2022) and total number of employees across their locations. 

LGU plant material has been licensed to `r length(unique(companies$licensee[companies$licensee_type == "company"]))` companies and `r length(unique(other_licensees$official_name))` public, non-profit, or individual licensees. The most active licensees are listed in the table below, all of whom have purchased over 30 licenses in the last 20 years. 


<!--- * Compare this number to the number of _unique_ companies in the licensing data (Hint: you need to reduce the number of rows in the licensing data so that you only have unique licensee 'names' or the type == company). --->
```{r, eval=F}
companies %>%
  count()
```

```{r}
top_licensees <- licenses %>%
  group_by(licensee) %>%
  count() %>%
  arrange(desc(n)) %>% 
  filter(n > 30)
DT::datatable(top_licensees, rownames = F,
              colnames = c("Licensee", "License count"))
```

## What do we know about company licensees?  

The companies licensing LGU plant varieties are primarily located in the United States with few international companies. The table below shows company headquarters' location by country, where 83.8% (973/1161) of companies are domestic.

```{r}
co_country <- companies %>%
  filter(!is.na(country)) %>% 
  group_by(country) %>%
  count() %>%
  arrange(-n)
DT::datatable(co_country, rownames = F,
              colnames = c("Company country HQ", "Count of licenses"))
```


<!--- * How 'big' are the companies? E.g. What is the range of sales data across all of the companies? What is the average sales number across all of the companies? --->

```{r, echo=FALSE, eval=FALSE}
# creating a highest employee number column because the employee numbers as presented in D&B are scattered across different columns (total, local, global)
companies <- companies %>%
  mutate(intermediate_employee_number = replace(total_employee_number, is.na(total_employee_number), local_employee_number[is.na(total_employee_number)]))

companies <- companies %>%
  mutate(highest_employee_number = replace(global_employee_number, is.na(global_employee_number), intermediate_employee_number[is.na(global_employee_number)]))
```

```{r, echo=FALSE, eval=FALSE}
# This was one attempt to group company size, but I think this was not used, right?
companies <- companies %>%
  mutate(size_range_employee = case_when(
    highest_employee_number > 2000 ~ "large", highest_employee_number > 1500 ~ "medium", 
    is.na(highest_employee_number) ~ NA_character_,
    TRUE ~ "small"))

companies %>%
  group_by(size_range_employee) %>%
  count() %>%
  arrange(-n)
```

Using the data from D&D Hoovers, we identified the the total sales per company as an estimate of company size. This sales value was calculated as the highest sales number per company level in order to capture company's size at the highest scale, rather than just one branch or region. These sales data can be used to categorize companies into categories: small and large. The U.S. Small Business Administration determines what is a small business based on its industry. The average annual sales for the farming industry to be considered a small business is less than 3 million dollars. (Note: Of the 263 companies (23%) that we could not identify in the D&B database, we assume that they are small because D&B preferences larger companies in its database).  

```{r, results = F}
companies <- companies %>%
  mutate(highest_sales_number = replace(global_sales_number, is.na(global_sales_number),                                   local_sales_number[is.na(global_sales_number)]))

# Doing something different. Based on what we know about the D&B data, we can actually assume that those we cannot identify in D&B are small, because D&B preferences large companies
# So I want to replace the NAs with small

companies <- companies %>%
  mutate(size_range_sales = case_when(
    highest_sales_number > 3000000 ~ "large",
    #is.na(highest_sales_number) ~ NA_character_,
    TRUE ~ "small"))

co_size_table <- companies %>%
  group_by(size_range_sales) %>%
  count() %>%
  arrange(-n)

co_sales_table <- companies %>%
  group_by(size_range_sales) %>%
  summarize(mean(highest_sales_number, na.rm = TRUE))
```

With this \$3 million division between small and large businesses, we find that `r co_size_table$n[co_size_table$size_range_sales == "small" & !is.na(co_size_table$size_range_sales)]` companies fall into the small category, where average small company makes \$685K dollars annually (2022 data). On the other hand, `r co_size_table$n[co_size_table$size_range_sales == "large" & !is.na(co_size_table$size_range_sales)]` companies fall into the large category, where average large company makes \$6.6B dollars annually.


```{r, results = F}
licenses_joined <- left_join(licenses, companies, by = "licensee")

licenses_joined %>%
  filter(licensee_type == "company") %>% 
  group_by(size_range_sales) %>%
  count()

licenses_joined %>%
  group_by(uni_state, size_range_sales) %>%
  count() %>%
  arrange(-n)

```

<!--- The two data frames can be joined by the licensee column. With size ranges, licenses can be categorized by university, licensee size, and FAO class. NA values of license size and FAO class are removed in order to accurately represent university action. --->

### Do certain LGUs trend towards small or large company clients?

On average, universities seem to license slightly more to smaller companies, with large companies accounting for about a third of licenses. We evaluate the relationship between plant variety licenses and the types of companies that purchase those licenses. Of the 6,966 licenses that we can connect to companies, 35% are to large companies. The percentage of small versus large companies that are licensees per university is quite similar across all states. For example, 29% of Washington's licenses are to large companies, though this number slightly increases: 34% of Texas's, 36% of Oregon's, 37% or North Carolina's, 38% of Iowa's and 43% of Other university's licenses are to large companies.  

DRAFT: Add in the percent of all companies that are large to identify if their licenzing activity is proportional to the number of companies. The result is large companies account for about 25% of companies but 30-40% of licenses, so they are slightly more active. 

```{r, eval = F}
licenses_joined %>%
  filter(licensee_type == "company") %>% 
  group_by(size_range_sales) %>%
  count() %>%
  ungroup() %>%
  mutate(percent = 100*n/sum(n)) 
```

```{r, out.width = "100%"}
geom_col1 <- licenses_joined %>%
  filter(licensee_type == "company") %>% 
  group_by(top_uni, size_range_sales) %>%
  count() %>%
  ungroup() %>%
  group_by(top_uni) %>%
  mutate(percent = round(100*n/sum(n))) %>%
  as.data.frame()

ggplot(geom_col1, aes(x = top_uni, y = n, fill = size_range_sales)) + 
  geom_col() +
  geom_text(aes(label = paste0(percent, "%")), position=position_stack(vjust=0.5),
            size = 3, color = "white") + 
  theme(axis.text.x = element_text(angle=70, vjust=0.5)) +
  scale_fill_viridis_d(option = "D", end = .6) +
  labs(title = "Size of companies licensed to per university",
    x = "", y = "Count", fill = "Company Size") +
  theme_minimal()
```

### Do certain crop categories trend towards small or large company clients?

There does not appear to be a clear trend whereby large companies are licensing specific types of crops. Large companies account for 42% of oilseed licenses, between 31-36% of vegetable, fruit and nut, and cereal licenses, though only 11% of roots and tubers. 

```{r, out.width = "100%"}
# Decided that to focus in on the question, you want to really know about how size relates to crop. University doesn't have too much to do with it right now.
licenses_joined %>% 
  filter(!is.na(size_range_sales),
         !is.na(crop_cat), crop_cat != "Ornamental") %>% 
  group_by(crop_cat, size_range_sales) %>% 
  count() %>% 
  ungroup() %>% 
  group_by(crop_cat) %>% 
  mutate(percent = paste0(round(100*n/sum(n)), "%")) %>% 
  arrange(desc(n)) %>% 
  mutate(percent = ifelse(crop_cat %in% c("Cereals",
                                          "Fruits & nuts",
                                          "Oilseeds",
                                          "Vegetables & melons",
                                          "Roots & tubers"),percent, "")) %>% 
  ggplot(aes(x = factor(crop_cat, level = unique(crop_cat)), 
             y = n, fill = size_range_sales)) + 
  geom_col() +
  geom_text(aes(label = percent), position=position_stack(vjust=0.5),
            size = 3, color = "white") + 
  theme(axis.text.x = element_text(angle=70, vjust=0.5)) +
  scale_fill_viridis_d(option = "D", end = .6) +
  labs(title = "Crops licensed by company size",
       x = "", y = "Count", fill = "Company Size") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


```{r, eval = F}
licenses_joined_NA_fao_class1 <- licenses_joined %>%
  filter(!is.na(fao_class1))

ggplot(licenses_joined_NA_fao_class1, aes(x = top_uni, fill = fao_class1)) + 
  geom_bar() +
  theme(axis.text.x = element_text(angle=70, vjust=0.5)) +
  scale_fill_viridis_d(option = "C") +
  labs(title = "Crops per University",
       x = "University", y = "License Count", fill = "FAO Class") +
  theme_classic()

licenses_joined_large <- licenses_joined %>%
  filter(size_range_sales == "large")

licenses_joined_large_NA_fao_class1 <- licenses_joined_large %>%
  filter(!is.na(fao_class1))

ggplot(licenses_joined_large_NA_fao_class1, aes(x = top_uni, fill = fao_class1)) + 
  geom_bar() +
  theme(axis.text.x = element_text(angle=70, vjust=0.5)) +
  scale_fill_viridis_d(option = "C") +
  labs(title = "Crops Licensed to Large Companies per University",
       x = "University", y = "License Count", fill = "Company Size") +
  theme_classic()
```


```{r, echo=FALSE, eval=FALSE}
licenses_joined %>%
  filter(uni_state == "Washington") %>%
  group_by(fao_class1) %>%
  count()

licenses_joined %>%
  filter(uni_state == "Washington", size_range_sales == "large") %>%
  group_by(fao_class1) %>%
  count()
```

<!--- ex) WRONG Washington works with 62% (1470/2378) small companies. 88% (1278/1470) of those small companies are interested in cereals. This is not significantly different than what Washington's normal cereals production at 89% (2122/2378).
    ### ex) Washington works with 6.6% (157/2378) large companies. 100% (157/157) of those large companies are interested in cereals. This is significantly different than what Washington's normal cereals production at 89% (2122/2378). --->

```{r, echo=FALSE, eval=FALSE}
# JOIN: 

# instead of finding the percentage of small/large companies in university licenses, this can determine how big/wealthy the companies are based on university. Also, this can determine how big/wealthy the companies are that work with different crops

licenses_joined %>%
  group_by(uni_state) %>%
  summarize(mean(highest_sales_number, na.rm = TRUE))

licenses_joined %>%
  group_by(fao_class1) %>%
  summarize(mean(highest_sales_number, na.rm = TRUE))

```


```{r, echo=FALSE, eval=FALSE}
licenses %>%
  group_by(licensee) %>%
  count() %>%
  arrange(-n)

licenses_wa_gen <- licenses %>%
  filter(licensee == "Washington Genetics")

licenses_wa_gen %>%
  group_by(effective_year) %>%
  count() %>%
  arrange(-n)

licenses_wscia <- licenses %>%
  filter(licensee == "Washington State Crop Improvement Association WSCIA")

licenses_wscia %>%
  group_by(effective_year) %>%
  count() %>%
  arrange(-n)

# Two Washington based licenses have made the most over the years. Washington Genetics has been making active licenses since 2015, peaking in 2016. WSCIA has been making active licenses since 2010, peaking in 2016. I will complete this action for multiple licensees.
```

```{r, echo=FALSE, eval=FALSE}
licenses %>%
 group_by(licensee) %>%
  count()

summary(companies$name)
# the total rows for licensee in each data frame do not match

# companies_unique <- companies %>% select(-link) %>% unique()
  ## this is the format if I wanted to filter out a column
  ## if I wanted to filter out a specific value, can use exclamation point

companies %>%
  group_by(licensee) %>%
  count() %>%
  filter(n>1)
# this yields 0 output, meaning every licensee in companies in unique

# licenses_joined <- left_join(licenses, companies, by = c("licensee"="official_name"))
  ## if ever wanted to join dfs by different column names
```
## DRAFT: Variety licensing networks 

The figures below display, by crop category, what operations are licensing material the most from LGUs. Circles are sized based on the number of license agreements made, and colored by the region in which the operation is based: West in dark purple, Northeast in dark blue, South in aqua, North Central in green, any companies outside of the US are in yellow, and non-companies in grey.  

```{r, results = F, echo = F}
df <- readRDS("~/Box/lgu/data_clean/df_final.RDS")
colors9 <- RColorBrewer::brewer.pal(name = "Spectral", n = 9)
library(igraph)
library(ggraph)
```

```{r, results = F, echo = F}
net_by_crop <- function(CROPCAT){
edges <- df %>% 
  filter(crop_cat == CROPCAT) %>% 
  select(uni_state, licensee) %>% 
  filter(!is.na(licensee)) %>% 
  group_by(uni_state, licensee)# %>% 
  #count() %>% 
  #rename(weight = n)
nodes_uni <- df %>% 
  select(uni_state, uni_region) %>% 
  unique() %>% 
  rename(node = uni_state, region = uni_region) %>% 
  mutate(type = 1)
nodes_licensee <- df %>% select(licensee, company_region) %>% unique() %>% 
  rename(node = licensee, region = company_region) %>% 
  mutate(type = 0)
nodes <- rbind(nodes_uni, nodes_licensee) %>% 
  filter(!is.na(node))
#net <- make_bipartite_graph(nodes$type, edges, directed = F)
net <- graph_from_data_frame(edges, vertices = nodes, directed = T)
V(net)$type <- bipartite_mapping(net)$type 

V(net)$deg <- log(igraph::degree(net, mode = "in")+1)+.5
maxlabel <- as.numeric(names(table(V(net)$deg)[length(table(V(net)$deg))-5]))
V(net)$label <- ifelse(V(net)$deg > maxlabel, V(net)$name, "")
V(net)$label.size <- 10
V(net)$color <- ifelse(V(net)$region == "West" , viridis::viridis(n = 5)[1],
                ifelse(V(net)$region == "Northeast", viridis::viridis(n = 5)[2],
                ifelse(V(net)$region == "North Central", viridis::viridis(n = 5)[3],
                ifelse(V(net)$region == "South", viridis::viridis(n = 5)[4],
                ifelse(V(net)$region == "International", viridis::viridis(n = 5)[5],      
                       "grey")))))
V(net)$label.color <- "darkgray"

V(net)$bw <- betweenness(net)
V(net)$comp <- igraph::components(net)$membership
main <- which(table(V(net)$comp) == max(table(V(net)$comp)))[[1]]
net <- induced_subgraph(net, V(net)$comp == main)
return(net)
}
```

```{r, echo = F}
net_cereal <- net_by_crop("Cereals")
net_fruit <- net_by_crop("Fruits & nuts")
net_oilseed <- net_by_crop("Oilseeds")
net_veg <- net_by_crop("Vegetables & melons")
net_root <- net_by_crop("Roots & tubers")
```

```{r, results = F, echo = F}
plot_net <- function(net, TITLE){
  ggraph(net, layout = "stress", bbox = 20) +
  geom_edge_link(width = 1, alpha = 0.5, color = "gray70") + 
  geom_node_point(size = V(net)$deg, color = V(net)$color, alpha = 0.5) + 
  geom_node_text(aes(label = V(net)$label),
                 size = 4,
                 color = "black") +
  theme_graph(base_family = "Arial") +
  labs(title = TITLE)
}
```

```{r, echo = F, out.width = "80%"}
plot_net(net_cereal, "Cereal crops")
plot_net(net_fruit, "Fruit and nut crops")
plot_net(net_veg, "Vegetable crops")
plot_net(net_oilseed, "Oilseed crops")
```


<!----
## Case study: University of California, Davis  
<!--- * Do a bit of exploring for UC Davis only (Hint: filter for just California). Within UC Davis, revisit the questions above to create a bit of a licensing 'profile' for UC Davis (essentially, tell a story about UC Davis licensing) 
  
University of California, Davis is the home institution of our research team, and is also among the top-10 active universities in our data with 210 licenses. Therefore, we take particular interest in California as a case study. Below, we see that California primarily produces fruits and nuts, and we note that all licenses are sold via exclusive licenses.

```{r}
licenses_ca <- licenses_joined %>%
  filter(uni_state == "California")

cacrop_table <- licenses_ca %>%
  group_by(crop_name_common) %>%
  count() %>%
  arrange(-n)

kable(cacrop_table, col.names = c("Crop category",
                                  "Count of licenses")) %>% kable_styling()
```

```{r, eval = F}
licenses_ca %>%
  group_by(agreement_type) %>%
  count() %>%
  arrange(-n)
```

Within California, we evaluated what companies they were licensing material to, particularly their size and if those companies were within the state of California. We find that UC Davis is mainly licensing to international companies, which is exceptional among LGUs, but that those international companies were normally small. Of the domestic companies they work with, very few are in the state of California, and nearly all of them are large (annual sales >\$3B).

```{r}
geo_table <- licenses_ca %>%
  mutate(instate = case_when(
    state == "California" ~ "in-state",
    country == "USA" ~ "in-country",
    T ~ "international"
  )) %>% 
  group_by(instate,size_range_sales) %>% 
  count()

kable(geo_table, col.names = c("Geographic location", "Company size",
                                  "Count of licenses")) %>% kable_styling()
```

--->
