---
title: "Who buys what?"
subtitle: "Plant material licensing for LGU Agricultural Colleges, 1980-2020"
author: ""
date: "`r Sys.Date()`"
output: html_document
---

```{r}
##### SET THIS! #####
path_to_box <- '~/Box/lgu/'
```

```{r setup, include = F}
knitr::opts_chunk$set(echo = T, warning = F, message = F, results = F)
knitr::opts_knit$set(root.dir = path_to_box)
```

```{r}
setwd(path_to_box)
licenses <- read.csv("data_clean/license_df.csv")
companies <- read.csv("data_clean/company_db_full.csv")
```

# Week 2-6: Describe, analyze, and visualize the data
## Explore all awards
* What are the columns in our data? Document the data by making a list of all the columns and writing our short descriptions of each of them
* How many licenses does each university sell each year? (Hint: group by uni_state and year, the count up their licenses (rows))
* Within UC Davis, how has the number of licenses changed over time?
  * Filter the data so that you have a data frame with just UC Davis
  * Now we can group by year to count up the licenses
* As you explore the data, take note: What issues do you see the with data? Is it 'clean'?

## Explore all the companies we have data for
* How many licensees do we know of based on our licensing data?
* Compare this number to the number of companies we have. (Hint: You will need to filter all_awards so that breeder = T). How well does the company data represent all of the licensees we have? 
  * Why might explain why some licensees are missing?
* Of the company daya that we have, where are they from? How big are they?
* Considering the licensing data, which companies have been most active across the different years? (Hint: with your breeder data, group by uni_state and year, the add up their funding amounts)
* Who are the most active (most licenses) companies across all years?

# Week 5-6: Analyze component data
## Summary statistics
* Across ALL funding, what is the average funding **per researcher** for each university. What is the standard deviation? What is the average funding **per project** for each university. What is the standard deviation?
* For just the BREEDERS, what is the average funding **per researcher** for each university. What is the standard deviation? What is the average funding **per project** for each university. What is the standard deviation?
* T-test: is one university better funded than another? 