---
title: "Who funds what?"
subtitle: "Research funding for LGU Agricultural Colleges, 2000-2020"
author: ""
date: "`r Sys.Date()`"
output: html_document
---

```{r}
##### SET THIS! #####
path_to_box <- '~/Box/lgu/'
```

```{r setup, include = F}
knitr::opts_chunk$set(echo = T, warning = F, message = F, results = F)
knitr::opts_knit$set(root.dir = path_to_box)
```

```{r}
setwd(path_to_box)
all_awards <- unique(read.csv('data_clean/awards_selected_states.csv'))
breeder_awards <- read.csv('data_clean/funds_df.csv')
inventors <- read.csv('data_clean/inventor_index.csv')
```


```{r}
all_awards <- breeder_awards %>% 
  select(-funding) %>% 
  mutate(breeder = T) %>% 
  right_join(all_awards) %>% 
  mutate(breeder = ifelse(is.na(breeder), F, T))
```


# Week 2-6: Describe, analyze, and visualize the data
## Explore all awards
* What are the columns in our data? Document the data by making a list of all the columns and writing our short descriptions of each of them
* How much money has each university gotten each year? (Hint: group by uni_state and year, the add up their funding amounts)
* Within UC Davis, how does funding vary across department?
  * Filter the data so that you have a data frame with just UC Davis
  * Now we can group by department to add up the funding
* As you explore the data, take note: What issues do you see the with data? Is it 'clean'?

## Explore all inventors and the inventors we have funding data for
* From our inventors data, how many inventors do we know of based on our licensing data?
* Compare this number to the number of inventors we have in our funding data. (Hint: You will need to filter all_awards so that breeder = T). How well does the funding data represent all of the inventors we have? 
  * Why might explain why some inventors are missing?
* Of the award data that we have for breeders, what departments are the breeders coming from? How many breeders are in each department?
* Now considering only the breeder award data, How much money has each university gotten each year? (Hint: with your breeder data, group by uni_state and year, the add up their funding amounts)
* Who are the top funders of awards to breeders?

# Week 5-6: Analyze component data
## Summary statistics
* Across ALL funding, what is the average funding **per researcher** for each university. What is the standard deviation? What is the average funding **per project** for each university. What is the standard deviation?
* For just the BREEDERS, what is the average funding **per researcher** for each university. What is the standard deviation? What is the average funding **per project** for each university. What is the standard deviation?
* T-test: is one university better funded than another? 

# Week 7: Link the data together
## Think about time

# Week 8: Analyze complete data
## Relationship between funding and innovation 
# Week 9-10: Write up final data